Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change

### 2020-12-15 Scripts/Manuscript.Rmd
William final comments

We projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd]. The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry. For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables . DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al. [-@Cheung:2010dt; -@Cheung:2016jd] for model details). Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al. *in revision*].

@@

I developed an index to evaluate range shifts in transboundary stocks under climate change (TI). This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2). The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., ^Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada. Therefore,

@@

where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005). I set a threshold of TI above which I consider a shared stock to have emerged from the historical internal variability of the TI. The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four), representing a probability of 68% that the transboundary index has emerged from historical variability). I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).

@@

Estimating the threat point of transboundary stock share

@@

First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100. I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2). Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability. The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db]. It is assumed to represent the historic baseline of shared distributions for each stock. I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SR{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR{tm}$). I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds]. The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$. This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{et} \geqq (SSR_{th} + \sigma)$ or $SSR_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al. (*in press*),

@@

Time of emergence of transboundary stocks

The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)). This means that on average, the distribution of transboundary stocks will diverge from their natural internal variability by 2029. In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100. Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100. About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g., Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010). The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020. After the mid-2020s, the rate of increase in the number EEZs slows from a 1% to an average of 0.7% per year. By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)). Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020. Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)). I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift). Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3).

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (\@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table @ref(tab:toeTableS1 for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (\@ref(fig:toeFigure2)A). There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (\@ref(fig:toeFigure2)A).

@@

I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)). Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods). I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001). Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table S2). On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)). Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).

@@

I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4))). My results suggest that 87% (n = 238) of the world EEZs experiences changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin see changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.

```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead). Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point. Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.s"}

@@

#### High present-day climate risk of on transboundary fisheries management

@@

Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds]. Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@@Pinsky:2018cb]. Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei]. I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change. My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management. While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; [@Gaines:2018sg; @PalaciosAbrantes:2020tv]. To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.

>>>>

I projected the distribution of marine species using a dynamic bioclimatic envelope model (hereafter called DBEM) [@Cheung:2010dt; @Cheung:2016jd]. The DBEM represents species’ physiology, habitat suitability, depth and latitudinal ranges, and spatial population dynamics as well as preferences for sea temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry. For pelagic species, the model uses environmental variables at the surface whereas demersal species distribution is driven by ocean bottom variables. DBEM then estimates species abundance and maximum catch potential (a proxy of maximum sustainable yield) over a $0.5^{\circ}$ x $0.5^{\circ}$ grid (see Cheung et al. [-@Cheung:2010dt; -@Cheung:2016jd] for model details). Importantly, DBEM is able to project catches for the world’s large marine ecosystems and changes in catches by Exclusive Economic Zones that are consistent with observational-based estimates of catch from 1950 to 2016 [@Cheung:2016jd, Cheung et al. *in revision*].

@@

I developed an index to evaluate range shifts in transboundary stocks under climate change (TI). This index is based on the shifts in the distribution centroid of a transboundary stock relative to the centroid of the neighboring EEZs that share this stock (Figure \@ref(fig:toeFigureS2)). The centroid of a transboundary stock was determined by the average (µ) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across the grid cells with the highest abundance within the neighboring EEZs sharing the stock e.g., Pacific halibut (*Hippoglossus stenolepis*) that is transboundary between the United States and Canada. Therefore,

@@

where $\mu_{TI}$ is the yearly mean TI ($\mu$) across all ten ensemble members in year *yr* and $SD_{TI}$ is the standard deviation of the smoothed (10-year moving average) TI across the ensemble members for the historical reference period (1951 - 2005). I set a threshold of TI above which I consider a shared stock to have emerged from the historical internal variability of the TI. The threshold was set as one standard deviation of the variability ($SD_{TI}$) (i.e., Equation \@ref(eq:four)), representing a probability of 68% that the transboundary index has emerged from historical variability). I tested the sensitivity of the calculated emergence index to  a higher emergence threshold of two times the standard deviation (i.e., representing a probability of 95% that the index has emerged).

@@

### Estimating the threat point of transboundary stock share

@@

First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100. I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2). Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability. The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db]. It is assumed to represent the historic baseline of shared distributions for each stock. I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$). I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds]. The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$. This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al. (*in press*),

@@

### Time of emergence of transboundary stocks

The average time of emergence of the transboundary index across the emerging stocks in all EEZs analyzed is projected to be 2029 $\pm$ 26 years (Figure \@ref(fig:toeFigure1)). This means that on average, the distribution of transboundary stocks will diverge from their natural internal variability by 2029. In total, the projected transboundary index of 5,745 stocks (63% of the studied stocks) will emerge from their historical variability between 2006 and 2100. Fish stocks with transboundary index that first emerged from their historical variability in each EEZ was projected to be as early as 2006 and the last one in 2100. About 55% (n = 3,154) of those emerging stocks are projected to have their transboundary index emerged between 2006 and 2020 (e.g., Peruvean anchoveta - *Engraulis ringens* - shared by Chile and Peru emerged in 2010). The number of EEZs with emerging stocks is projected to increase steeply from 87% in 2006 to 96% in 2020. After the mid-2020s, the rate of increase in the number EEZs slows from a 1% to an average of 0.7% per year. By 2081, all of the EEZs analyzed here will have at least one emerging stock (Figure \@ref(fig:toeFigure1)). Moreover, 96% of the world’s EEZs saw at least one stock having their distribution shifts between 2006 and 2020 and around one fifth (19%) of the EEZs have an average time of emergence of the transboundary index across stocks prior to 2020. Shifts in stocks’ distributions will continue to exceed historical limits steadily towards the end of the $21^{st}$ century (Figure \@ref(fig:toeFigure1)). I based the transboundary index on a transboundary stock’s centroid defined as the average grid cells that present the higher stock’s abundance (see Methods - Calculating an index of transboundary range shift). Results from the sensitivity analysis suggest that this method of estimating the time of emergence of the transboundary index is robust to the abundance level (e.g., the top 95%) of the stock (Figure \@ref(fig:toeFigureS3)).

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A). There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A).

@@

I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)). Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods). I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001). Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)). On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)). Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).

@@

I have identified the stocks that are projected to have changes in stock share ratio beyond an EEZ’s threat point and estimated the changes in the stock share ratio by the early (2020-2040) and the mid $21^{st}$ century (2040-2060), relative to the recent past (1951-2005) (see methods – Equation \@ref(eq:five)). By the early $21^{st}$ century, the global average proportion of transboundary stocks that are projected to change their stock share ratio beyond the EEZ’s threat point is 18% $\pm$ 15% (Figure \@ref(fig:toeFigure5)A). The changes in stock share ratio are expected to increase slightly towards the mid $21^{st}$ century (Figures \@ref(fig:toeFigureS4) and \@ref(fig:toeFigureS5)). However, changes in stock share ratio for specific stocks and in some EEZs are projected to be exceptionally high. For example, Guatemala shares *Panulirus gracilis* stock with Mexico, and Guatemala’s share is projected to increase by >50%, that is from 12% to 26% by the early $21^{st}$ century. Towards the mid $21^{st}$ century, changes in stock share ratio per EEZ will not be substantial (Figure \@ref(fig:toeFigureS5)) relative to the early $21^{st}$ century but are expected to increase in intensity (Figure \@ref(fig:toeFigureS4)). My results suggest that 87% (n = 238) of the world EEZs experiences changes in stock share ratio beyond a country’s threat point by the early $21^{st}$ century (Figure \@ref(fig:toeFigure5)A). These EEZs see one third (33% $\pm$ 21%) of their transboundary stocks changing the stock share ratio beyond the threat point (Figure \@ref(fig:toeFigure5)A). Particularly, Brazil, French Guiana, Guam, Kergelen and Pitcairin see changes beyond the threat point for all of the transboundary stocks analyzed, while Venezuela (6%), Bonaire (9%), China (10%) and Vietnam (10%) are the only countries with less than 10% of stocks changing beyond the threat point.

```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead). Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point. Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}

@@

### High present-day climate risk of on transboundary fisheries management

@@

Global society has set the ambitious goal to manage all fisheries sustainably (SDG 14 – Life below water) by 2030; achieving this goal would have clear effects on multiple others [@UnitedNations2018; @Singh:2017ds]. Preparing anticipatory policy to species on the move is key to achieve these sustainable development goals and effective governance of the world’s fisheries [@Pinsky:2018cb]. Here, I have addressed two fundamental steps to achieve sustainable international fisheries in a changing world [@Link:2010ei]. I have identified the transboundary stocks that will see a significant shift compared to their historical average in shared distribution, the year in which their distribution will undergo such a shift, as well as the degree of change. My results have direct implications for ocean governance as the data provided can inform and be used by regional management bodies and countries to anticipate the potential complications brought about by climate change to international fisheries management. While future studies, specifically at more localized scale, will provide valuable nuance in designing effective policies, my results provide an important baseline on which to build when preparing ocean governance for species on the move [@Pinsky:2018cb; @Gaines:2018sg; @PalaciosAbrantes:2020tv]. To address the issues raised in this chapter, international fisheries management will require coordination between nations to generate adaptive, collaborative and equitable management strategies to address the uncertainties of a changing world.

### 2020-12-15 Scripts/Supplements.Rmd
William final comments

```{r toeFigureS1, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}

@@

```{r toeFigureS2, eval = T,  echo = F, fig.width=9, fig.height=4, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}

@@

read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
  select(-statistic,
         "Habitat preference comparrison" = comparrison,
         "Significan level" = signif.level,
         "Observation difference" = dif.com.obs.dif,
         "Critical difference" = dif.com.critical.dif,
          "K-W Difference" = dif.com.difference) %>% 
  kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
  # kable("latex",
        # booktabs = T,
        # caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  # ) %>% 
  # kable_styling(full_width = F,
                # position = "center")

@@

read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
  select(-statistic,
         "Sub region comparrison" = comparrison,
         "Significan level" = signif.level,
         "Observation difference" = dif.com.obs.dif,
         "Critical difference" = dif.com.critical.dif,
          "K-W Difference" = dif.com.difference) %>% 
  kable(
        booktabs = T,
        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  )  
  # kable("latex",
  #       booktabs = T,
  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  # ) %>% 
  # kable_styling(full_width = F,
  #               position = "center")

>>>>

```{r toeFigureS1, eval = T,  echo = F, fig.width=12, fig.height=7, fig.cap =  "World Exclusive Economic Zones used in this study as defined by the Sea Around Us and their centroids (points)."}

@@

```{r toeFigureS2, eval = T,  echo = F, fig.width=12, fig.height=5, fig.cap = "Graphical representation of the transboundary index (TI) to determine the time of emergence of transboundary stocks. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighbouring Exclusive Economic Zones sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overpasses historical values (Bottom panel)."}

@@

read.csv("~/GitHub/TransEmergence/Tables/kw_regional.csv") %>% 
  select(-statistic,
         "Habitat preference comparrison" = comparrison,
         "Significan level" = signif.level,
         "Observation difference" = dif.com.obs.dif,
         "Critical difference" = dif.com.critical.dif,
          "K-W Difference" = dif.com.difference) %>% 
  # kable(caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions")
  kable("latex",
        booktabs = T,
        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  ) %>%
  kable_styling(full_width = T,
                position = "center")

@@

read.csv("~/GitHub/TransEmergence/Tables/kw_habitat.csv") %>% 
  select(-statistic,
         "Sub region comparrison" = comparrison,
         "Significan level" = signif.level,
         "Observation difference" = dif.com.obs.dif,
         "Critical difference" = dif.com.critical.dif,
          "K-W Difference" = dif.com.difference) %>% 
  # kable(
  #       booktabs = T,
  #       caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  # )  
  kable("latex",
        booktabs = T,
        caption = "Statistical results for multiple comparison test after Kruskal-Wallis on time of emergence across United Nations sub regions"
  ) %>%
  kable_styling(full_width = F,
                position = "center")
